---
title: "Data Visualisation with ggplot2 - Project"
output: html_notebook
---

# AfterWork Data Science: Data Visualisation with Python Project

## 1. Defining the Question

### a) Specifying the Data Analysis Question
Specify the reasearch question that you'll be answering. i.e. Provide strategy recommendations that will lead to revenue growth.

### b) Defining the Metric for Success
The solutions to the following questions will help us answer our research question:

* When is the best time of year to book a hotel room?
* When is the optimal length of stay in order to get the best daily rate?
* How will you know if a hotel was likely to receive a disproportionately high number of special requests?

### c) Understanding the context
Provide some background information....

### d) Recording the Experimental Design
Describe the steps/approach that you will use to answer the given question.

### e) Data Relevance
How relevant was the provided data?


## 2. Reading the Data

```{r}
# Load the data below
install.packages("tidyverse")
library(tidyverse)
# --- 
# Dataset url =  https://bit.ly/2WvQbhJ
# --- 
# YOUR CODE GOES BELOW
hotels_df <- read_csv("hotel_bookings.csv")
# 
```

```{r}
# Checking the first 5 rows of data
# ---
# YOUR CODE GOES BELOW
head(hotels_df, 5)
# 
```

```{r}
# Checking the last 5 rows of data
# ---
# YOUR CODE GOES BELOW
tail(hotels_df, 3)
# 
```

```{r}
# Sample 10 rows of data
# ---
# YOUR CODE GOES BELOW
sample_n(hotels_df, 10)
# 
```

```{r}
# Checking number of rows and columns
# ---
# YOUR CODE GOES BELOW
dim(hotels_df)
#  
```

```{r}
# Checking datatypes
# ---
# YOUR CODE GOES BELOW
glimpse(hotels_df)
# 
```

```{r}

```

Record your general observations below:

Observation 1
Observation 2


## 3. External Data Source Validation
The data is originally from the article Hotel Booking Demand Datasets, by Nuno Antonio, Ana Almeida, and Luis Nunes for Data in Brief, Volume 22, February 2019.


## 4. Data Preparation
### Performing Data Cleaning
```{r}
# Checking datatypes and missing entries of all the variables
# ---
# YOUR CODE GOES BELOW
colSums(is.na(hotels_df))
# 

```

We observe the following from our dataset:

Observation 1
Observation 2

```{r}
# Checking how many duplicate rows are there in the data
# ---
# YOUR CODE GOES BELOW
hotels_df[duplicated(hotels_df), ]
# 
```

We choose to keep the duplicates because we don't have a unique identifier to actually proof that we have duplicates.

```{r}
# Checking if any of the columns are all null
# ---
# YOUR CODE GOES BELOW
colSums(is.na(hotels_df))

# 
```

We observe the following from our dataset:

Observation 1

```{r}
# Checking if any of the rows are all null
# ---
# YOUR CODE GOES BELOW
# 
```

We observe the following from our dataset:

Observation 1
Observation 2

```{r}
# Checking the correlation of the features through the use of 
# visualizations the correlation using heatmap
# ---
# YOUR CODE GOES BELOW
# 
```

We observe the following from our dataset:

Observation 1
Observation 2

```{r}
# Dropping company column because it has alot of missing values 
# and we won't need to answer any of our questions
# ---
# YOUR CODE GOES BELOW
hotels_df = select(hotels_df, -c("company"))
head(hotels_df)
# 

```

From the data variable description we see that the Distribution Channel categoricy that tells us about Booking distribution.

The term “TA” means “Travel Agents”
The term “TO” means “Tour Operators”
This allows us to fill the missing values in the agents column with TO

```{r}
# We replace the mising values i.e. for TO
# ---
# YOUR GOES BELOW
hotels_df$distribution_channel[is.na(hotels_df$distribution_channel)] <- TO
#
```

```{r}
# We drop rows where there is no adult, baby and child as 
# these records won't help us.
# ---
# YOUR GOES BELOW
na.omit(hotels_df, children,adults,babies)
#
```

```{r}
# We replace missing children values with rounded mean value
# ---
# Hint i.e. use round()
# ---
# YOUR GOES BELOW
#
```

```{r}
# Checking for missing values in the dataframe
# ---
# YOUR GOES BELOW
is.na(hotels_df)
#
```

```{r}
# Converting the datatypes of the following columns from float to integer
# i.e. children, company, agent
# ---
# YOUR GOES BELOW
# 
```


## 5. Solution Implementation

### 5.a) Questions

```{r}
# 1. How many bookings were cancelled?
# ---
# Visualisation: Barplot
select(hotels_df, reserved_room_type)
head(hotels_df)
```


```{r}
# 2. What was the booking ratio between resort hotel and city hotel?
# ---
# Barplot of booking ratio between resort hotel and city hotel
```


```{r}
# 3. What was the percentage of booking for each year?
hotels_df %>%
    group_by(arrival_date_month) %>%
    summarise(percent = arrival_date_month/sum(arrival_date_month))
#select(hotels_df, arrival_date_month)
# ---
# 
```

```{r}
# 4. Which were the most busiest months for hotels?

hotels_df %>%
    group_by(arrival_date_month) %>%
    summarise(count=n())
#select(hotels_df, arrival_date_month)
# ---
# 

```


```{r}
# 5. From which top 3 countries did most guests come from?

hotels_df %>%
    group_by(country) %>%
    summarise(count=n())
# --- 
# YOUR GOES BELOW

#

```

```{r}
# 6.a) How long do most people stay in hotels?
#   b) By city and resort? Separate the data by hotel
hotels_df %>%
  group_by(hotel, country) %>%
  summarize(stays_in_week_nights = sum(stays_in_week_nights))
# ---
# 
```


```{r}
# 7. Which was the most booked accommodation type (Single, Couple, Family)?

hotels_df %>%
    group_by(reserved_room_type) %>%
    summarise(count=n())
select(hotels_df, stays_in_week_nights)
# ---
#
```

### 5.b) Recommendations
From the above analysis, below are our recommendations:



## 6. Challenging your Solution
In this step, we review our solution and implement approaches that could potentially provide a better outcome. In our case, we could propose the following question that wasn't answered in our solution because it couldn't have greatly contributed to our recommendation.

```{r}
# When should hotels provide special offers?
# ---
# YOUR GOES BELOW
# 
```

Our observations:
-

How does this observation tie to our solution?

## 7. Follow up questions
During this step, you rethink and propose other ways that you can improve your solution.

a). Did we have the right data?
b). Do we need other data to answer our question?
c). Did we have the right question?


